Extracting Function Signatures from Program Traces

### 2020-05-14 paper/main.tex
abstract

\begin{abstract}
  Retrofitting a type system onto an existing dynamic language which was not
  designed to be statically typed is challenging. This paper takes the first
  step in the direction of designing a type system for the R programming
  language. We argue that R is particularly challenging due to its unique
  combination of features and to the programming idioms that have developped
  over the last two decades. As our starting point, we infer type signatures
  for \NUMFUNCTIONS functions from \NUMPACKAGES packages amongst the most
  widely used open source R libraries. The signatures use a minimal type
  language consisting of basic types, vectors, classes and disjunctions. We
  purposefully leave out notions such as subtyping, higher-order types, and
  polymorphism. Our goal is to understand what benefits a simple lanugage
  can yield.  We also implement a contract system that can use the inferred
  types to validate their correctness. We use the reverse depedencies of the
  packages to evaluate the benefits of typing.
\end{abstract}
\maketitle

>>>>

\begin{abstract}
The R programming language is widely used in a variety of domains for tasks
related to data science. R was designed to favor an interactive style of
programming with minimal syntactic and conceptual overhead. This design is
well suited to interactive data analysis, but a bad firt for tools such as
compilers or program analyzers which must generate performant code or catch
programming errors.  In particular, R has no type annotations, its
operations are dynamically checked at runtime. The starting point for our
work are the twin questions, \emph{what expressive power is needed to
  accurately type R code?} and \emph{which type system is the R community
  willing to adopt?} Both questions are difficult to answer without actually
designing a type system and attempting to convince users and package
developers to try the proposed system.  The goal of this paper is to provide
data that can feed into the design process. To this end, we perform a large
corpus analysis with aim to gain insights in the degree of polymorphism
exhibited by idiomatic R code and explore potential benefits that the R
community could accrue from even a simple type system.  As a starting point,
we infer type signatures for \NUMFUNCTIONS functions from \NUMPACKAGES
packages among the most widely used open source R libraries. The signatures
use a minimal type language with only basic types, vectors and classes.  We
implement a contract checker that can be used to validate the inferred type
signatures.
\end{abstract}
\maketitle
